% Created 2020-10-19 Mon 16:14
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{unicode-math}
\usepackage{unicode}
\author{Cynthia Liu}
\date{\today}
\title{Computer Science 3MI3 – 2020 Assignment 1: A language of expressions}
\hypersetup{
 pdfauthor={Cynthia Liu},
 pdftitle={Computer Science 3MI3 – 2020 Assignment 1: A language of expressions},
 pdfkeywords={},
 pdfsubject={A demonstration of how to represent tree-like data in Prolog.},
 pdfcreator={Emacs 27.1 (Org mode 9.3)}, 
 pdflang={English}}
\begin{document}

\maketitle
\tableofcontents


\section{Introduction}
\label{sec:org9fae899}
This is the documentation for Assignment 1 for COMPSCI3MI3 2020fall. 
It is about an expression language consisting of integer constants and seven prefix operators, 
like negation, absolute value, plus, times, minus, and exponent. 
Also, base on that, implementing variables and substitution, as well as boolean expressions.
This Assignment is written in both Scala and Prolog. 

\section{Part one}
\label{sec:orgea006d3}
\subsection{Representation}
\label{sec:org72e2954}
The integer constants and operators are expressed in the new type \texttt{Expr}.
\begin{itemize}
\item The constructor Const take an \texttt{Int} as argument.
\item The constructor Neg and Abs take an \texttt{Expr} as argument.
\item The constructor Minus, Plus, Times and Exp take two \texttt{Expr} as arguments.
\end{itemize}
\begin{verbatim}
sealed trait Expr
case class Const[A](value:Int) extends Expr
case class Neg[A](value:Expr) extends Expr
case class Abs[A](value:Expr) extends Expr
case class Minus[A](value1:Expr, value2:Expr) extends Expr
case class Plus[A](Tvalue1:Expr, value2:Expr) extends Expr
case class Times[A](value1:Expr, value2:Expr) extends Expr
case class Exp[A](value1:Expr, value2:Expr) extends Expr
\end{verbatim}



\subsection{Interpreter}
\label{sec:org8a6a204}

\section{Part two}
\label{sec:org7c8c2da}
hgfd$\backslash$

\section{Part three}
\label{sec:orge841626}
gsdajg

\section{Part four}
\label{sec:org7feb859}
bfdsd
\end{document}
